---
title: "Analysis Workflow"
author: "Research Bow"
date: "`r Sys.Date()`"
output: 
  word_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#install.packages(c("gtable","gtsummary", "flextable", "tidyverse",
"correlation","gt", "psych", "rstatix", "flextable","jtools","modelsummary"  ))


library(tidyverse) # data tidying
library(correlation) # prepares nice correla
library(gtable) # Allows for custom tidy tables, which render to Latex, word, pdf, html
library(gt) # More control on making tables
library(psych) # Useful packages for working with psychometrics and making summary tables
library(rstatix) # Useful tools for working statistical analyses
library(gtsummary) # Has ready made tables for summary stats
library(flextable) # Converts regression tables to pretty publication tables. May not work on Mac.
library(jtools) # Useful for generating APA ready plots
library(modelsummary) # Mac friendly alternative for regression tables. 
```

Before cracking on with the second part of today's session, we are going to quickly learn about how R can be used to render documents. This is done using the `Knit` button. To tidy our end document, we will also use some extra commands in our code chunks. The `echo=FALSE` is used to prevent our code from being rendered. Meanwhile, `message = FALSE` is used to prevent any warning messages from being rendered.

```{r, echo=FALSE, message = FALSE, results= FALSE}
#code to upload our newly saved file into R.
df_tidy <- read_csv(url("https://raw.github.com/Rhys-Maredudd/Research-Bow-R-Workshop/main/~tidy_data.csv")) %>% mutate(group = as.factor(group))
summary(df_tidy) 
```

## Running and reporting analyses

Now that the hard work of preparing our data is complete. We can run our analyses! Today we will perform t-tests to compare the effect of the intervention on depression, anxiety and stress.

From there, we will perform a regression analysis to predict depression scores whilst controlling for exposure to stressful life events, social support, and the intervention.

We will also use this to demonstrate how we can use R to generate our analysis reports. So no more copy and pasting. Just smoooooooth sailing.

Here we will be taking advantage of the `flextable` and `gtsummary` packages to conduct and report our statistics. These package makes it super easy to generate publication ready tables, and it can be rendered into Word documents thorugh R!

For more information on how to use the package, please check out the `flextable` [vignettes](https://ardata-fr.github.io/flextable-book/) and the `gtsummary` [vignettes](https://www.danieldsjoberg.com/gtsummary/articles/tbl_summary.html).

The code might look a little intimidating, but we will walk you through it so that you can also use this for your own analysis.

For each analysis, we will display the raw version and the tidied version. We will also show you how to render your analysis to finish the session.

## Summary Statistics

The code for the summary statistics table has been set to display most of the common descriptive summary statistics. Any columns that are not needed can be deleted manually in the rendered word document (Or by digging into `get_summary_stats()` by looking up the [R documentation](https://www.rdocumentation.org/packages/rstatix/versions/0.7.2/topics/get_summary_stats)).

```{r, echo = FALSE, message= FALSE}



 summary_stats <- df_tidy %>%
  select(stress, anxiety,  depression, # selecting our required variables
         eventsscore , SOsupport  ) %>%
  psych::describe() %>%
  as.tibble(rownames = "rowname") %>% 
   select(variable = rowname, #changing name of first column
          mean, sd, median) %>% # select the summary stats you need
  mutate(mean = round(mean,3), # rounding decimal places
         sd = round(sd, 3),
         median = round(median, 3)
  )
 
 summary_stats %>% gt()

 # Alternative method, quicker for coding, but requires manual deleting in rendered word document. 
 
table <- df_tidy %>%  
  get_summary_stats(
    stress,anxiety,  depression, eventsscore , SOsupport,  # columns to calculate for
    type = "common") %>%               # all common summary stats to return
    as.data.frame()


gt(table)  # Converts table to gt, which makes it prettier and allows for rendering into Word. 
 
```

Using R also allows us to report our summary stats in text! This can save us time, and prevent typing errors. It does require some confidence in our coding, but we can always test it in the code chunks above if needed. The important part is to create our data objects first, we can easily retrieve it for the document. In this case, we will use our `summary_stats` object to retrieve our stats.

So, if say we wanted to report the overall mean and sd values for stress, we would use the following code `r summary_stats %>% filter(variable == "stress") %>% select(mean)` (*sd* = `r summary_stats %>% filter(variable == "stress") %>% select(sd)`).

#### Task

Report the mean and sd values in text for the depression and anxiety subscales.

### Cronbach's alphas

This is probably the most complicated part of the worksheet, as it requires using a combination of different functions. However, we have tried to compile the code so that you can copy, paste and adjust to your own needs.

```{r, echo = FALSE, message = FALSE}

# Compiling alphas

## Step 1: compiling subscale variables
stress_alpha <- df_tidy %>% dplyr::select(dass1, dass6, dass8,
                                       dass11, dass12, dass14, dass18)
anxiety_alpha <- df_tidy %>% dplyr::select(dass2, dass4, dass7,
                                        dass9, dass15, dass19, dass20) 
depression_alpha <- df_tidy %>% dplyr::select(dass3, dass5, dass10, 
                                           dass13, dass16, dass17, dass21) 

## Step 2: Running the cronbach.alpha test, and collecting only the score.
## If the commands below are not working, run `instal.packages("ltm")` in the console first. 

str_alpha_value <- ltm::cronbach.alpha(stress_alpha)[[1]]
anx_alpha_value <- ltm::cronbach.alpha(anxiety_alpha)[[1]]
dep_alpha_value <- ltm::cronbach.alpha(depression_alpha)[[1]]


## Step 3:Creating Subscale names and cronbach alpha values objects - make sure values align with names

alpha_values <- c(str_alpha_value , anx_alpha_value, dep_alpha_value)

Measure_name <- c("DASS Stress", 
              "DASS Anxiety",
              "DASS Depression")

## Step 4: Creating a pretty table
data.frame(Measure_name, alpha_values) %>% # Choosing our objects from step 3
  mutate(alpha_values = round(alpha_values, 2)) %>% # Rounding to 2 decimal places
  dplyr::rename( `Measure` = Measure_name, # Renaming so we can include spaces
    `Cronbach alpha` = alpha_values ) %>% 
  gt() # Using gt() to make prettier table that renders to word.
```

## T-test report

### Raw results

```{r, echo=FALSE, message = FALSE}
result <- t_test(stress ~ group, data = df_tidy)
#result

summary_stats_t <- df_tidy %>% 
                      group_by(group) %>%
                      summarise(mean_stress = round(mean(stress),3),
                                sd_stress = round(sd(stress), 3))

#summary_stats_t 
```

Welch's t-test demonstrated that stress in the control group (*M* = `r summary_stats_t %>% filter(group == "control") %>% select(mean_stress)`, *sd* = `r summary_stats_t %>% filter(group == "control") %>% select(sd_stress)`) was significantly higher (*t*(`r result %>% select(df) %>% print(digits = 3)`) = `r result %>% select(statistic) %>% print(digits = 3)`, *p* = `r result %>% select(p) %>% print(digits = 3)`) than the experiment group (*M* = `r summary_stats_t %>% filter(group == "experiment") %>% select(mean_stress)`, *sd* = `r summary_stats_t %>% filter(group == "experiment") %>% select(sd_stress)`).

### Publication ready(er) results

```{r, echo=FALSE, message = FALSE}
  df_tidy %>%
  select(group, anxiety, stress, depression) %>%
  tbl_summary(by = group, 
              missing = "no",
              statistic = list(all_continuous() ~ "{mean} ({sd})")) %>%
  add_p(everything() ~ "t.test") %>%
  modify_header( # header needs to be modified to display test statistc and df 
    statistic = "**t-statistic**", 
    parameter = "**df**"
  )  %>%
  # add a header to the statistic column, which is hidden by default
  modify_fmt_fun(c(statistic , parameter) ~ style_sigfig)
```

## Corellation analysis

### Rough results

```{r, results = "asis", echo = FALSE, message = FALSE}

cor_data <- df_tidy %>% 
  select( anxiety, stress, depression, eventsscore , SOsupport)

results <- correlation(cor_data)

summary(results, redundant = TRUE)

```

### Publication ready(er) results

Please note that the correlation matrix is left with redundant values. This is to allow you to decide if you want to manually display the lower or upper triangle. Also, a bug in the `correlation` package rearranges the display in a displeasing manner when `redundant = FALSE`. But thankfully, we can edit our output to our needs in the rendered word document.

```{r, results = "asis", echo = FALSE, message = FALSE}

 df_tidy %>% dplyr::select( anxiety, stress, depression, eventsscore , SOsupport) %>%
    rename(`Anxiety` = anxiety ,
           `Stress` = stress,
           `Depression` = depression,
           `Stressful Life Events` =  eventsscore,
           `Significant Other Support`=  SOsupport) %>% 
  correlation::correlation() %>% 
  summary(redundant = TRUE) %>%
  display()

```

## Regression analysis

```{r, echo = FALSE, message = FALSE}
# Setting the model
model <- lm(scale(depression) ~ group + scale(eventsscore) + scale(SOsupport), df_tidy)
## Wrapping variables in scale() to get standardised Beta's. Remove the scale() if regular coefficents are needed.
```

### Raw results

```{r, echo = FALSE}
# Ugly (but useful) results 
#summary(model)
regression_coef <- tidy(model)

regression_model_fit <- glance(model)

# -   -   SOsupport is a significantly and negatively associated with Depression ($\beta$ = `r round(regression_coef) %>% filter(term == "SOsupport)") %>% select(estimate), 2)`, *p* = `r round(regression_coef %>% filter(term == "SOsupport") %>% select(p.value), 2)`).
# -   SOsupport is a significantly and negatively associated with Depression ($\beta$ = `r round(regression_coef %>% filter(term == "SOsupport") %>% select(estimate), 2)`, *p* = `r round(regression_coef %>% filter(term == "SOsupport") %>% select(p.value), 2)`).
```

We can also extract our regression analysis into text, by using the `broom` package to extract our model summary data. We use the `tidy()` function to extract the coefficent values. And we use the `glance()` to extract the model fit statistics. This is very useful to save time! But can look intimidating. The best tip is to test the code in the code boxes, then copy and paste it into the in-line code. Below are some examples of the model results:

-   SOsupport is a significantly and negatively associated with Depression ($\beta$ = `r round(regression_coef %>% filter(term == "scale(SOsupport)") %>% select(estimate), 2)`, *p* = `r round(regression_coef %>% filter(term == "scale(SOsupport)") %>% select(p.value), 3)`).
-   Model fit is significant (*F* (`r regression_model_fit$df`, `r regression_model_fit$df.residual`) = `r round(regression_model_fit$statistic, 2)`, *p* = `r  plyr::round_any(regression_model_fit$p.value , .001, ceiling)`), with `r round(regression_model_fit$adj.r.squared * 100, 2)`% of model variance explained by the $R^2_{adj}$ .

### Publication ready(er) results

I have recently discovered the `flextable` package, and it is amazing! It will convert your model into a publication ready format in word with minimal effort. Simply take your defined model, and wrap it in the `as_flextable()` function. More information on `flextable` can be found [here](https://ardata-fr.github.io/flextable-book/index.html).

For mac-users, there appear to be some headaches with the `flextable` package. So here's a mac-friendly version that makes use of the `modelsummary` package. More information on using the modelsummary  [website](https://modelsummary.com/vignettes/modelsummary.html).

```{r, echo = FALSE, message = FALSE}
# Pretty results 
flextable::as_flextable(model) # this approach may not work for mac users. Comment it out if it misbehaves.

modelsummary(model,
             shape = term ~ model + statistic , # allows for nicer table format
             statistic = c('std.error', "p.value"), #include the desired summary statistics
             gof_omit = "AIC|BIC|Log.Lik.|RMSE", # remove unwanted model fit statistics
             output = "gt") # choose printing style


```

### Publication ready plot

And what use is a fancy table without a fancier plot?! APA guidelines are a little hazy when it comes to plot standards, but here is a good estimate. The title, subtitle and caption have been left empty here, as sometimes it is more useful to edit this within word (sometimes...).

```{r, echo = FALSE, message = FALSE}
#Generating plot
 plot <- ggplot(df_tidy,
                aes(x = SOsupport, y = depression, col = group)
                ) +
  geom_point(alpha = .7, position = "jitter")+
  geom_smooth(method = "lm", se = FALSE) 
  

#Adding labels and APA theme
  plot +
    labs(x = "Support from significant others",
         y = "Depression score (DASS)",
         col = "Condition") +
     theme_apa(legend.pos = "right", 
               legend.use.title = TRUE) 
    

```

## Finishing session

Now click on `Knit` to prepare your word document with all the prepared analyses.
